Dark Control:\\ A Unified Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-04-09 paper/paper.tex
monster pass

\begin{abstract}
The default mode network (DMN) is believed to subserve the human
baseline mental activity.
%
Many research streams converge to its likely role in evolutionarily adaptive
envisioning of mental scenarios to simulate the future.
Such functional roles become more likely due to
this brain network's highest energy consumption in the brain and
its intimate coupling with conscious awareness
\textbf{(XXX Guillaume: It escaladed quickly.)}.
%
The present concept paper proposes a \textit{process model}
in trying to explain \textit{how} the DMN may actually implement continuous
assessment and prediction of the environment to guide action choices.
DMN function is recast in the mathematical terms of
control theory and reinforcement learning based on Markov decision processes.
We argue that our formal account of DMN function
naturally accommodates proposed cognitive accounts as special cases and
parsimoneously explains
experimental findings in animals and humans.
%
Formal process models of the neural implemention
of future predictions probably realized in the DMN
could offer insight into the statistical mechanisms underlying
human behavior.
\end{abstract}

@@

\section{Introduction}
%
When left unperturbed, the human brain is not at rest.
Rather, the brain continues to metabolize large quantities of
oxygen and glucose energy to maintain
potentially purposeful neuronal computation
in the absence of a focused behavioral goal
\citep{kenet2003spontaneously, fiser2004small}.
The baseline energy demand is subject to surprisingly little modulations
when mentally processing environmental challenges
\citep{raichle2001pnas}. \textbf{(XXX Guillaume: What about the energy consumption compare to sleep? Danilo: I'd love to cite that
but I am not aware of such studies)}
This dark matter of brain physiology begs the question
of underlying biological purpose
\citep{raichle2006brain}.
What has early been described as the "stream of consciousness"
in psychology \citep{james1890principles}
found a potential neurobiological manifestation
in the so-called "default mode network" (DMN)
\citep{shul1997}. This networked collection of some of the highest regions in
the human association cortex \citep{mesulam1998sensation, margulies2016situating}
consistently
increased in neural activity during
unfocused everyday mind wandering \citep{raichle2001pnas}.
The DMN may exert higher-order control on behavior
as the human brain was argued to features "a hierarchy of brain systems with
the DMN at the top and the salience and dorsal attention systems
at intermediate levels, above thalamic and unimodal sensory
cortex" \citep{carhart2010default}.
In the beginning of the 21st century,
brain imaging may have been first technique
to allow for the discovery of a unique brain network that subserves
baseline mental activities
\citep{raichle2001pnas, bzdok2015resting}.

\subsection{Default mode network: higher-order control of the organism}
The human DMN accounts for a large part of the brain's base line activity
and typically decreases in activity during psychological experiments.
The standard mode of
information maintenance and manipulation in the brain has
been argued to mediate evolutionarily conserved functions
\citep{brown1914nature, binder1999conceptual, buzsaki2006rhythms}.
This is all the more likely because the DMN contains hotspots
of highest energy consumption in the entire central nervous system
\citep{raichle2001pnas}.
DMN activity also persists to a substantial degree during
sleep and under anesthesia \citep{randy2008}.
Today, many authors believe that the DMN implements
probabilistic estimation of past, hypothetical, and
future events
\citep{fox2005, hassabis2007patients, schacter2007remembering, binder2009,
randy2008}.
This brain network
might have emerged to continuously predict environmental events using
mental imagery as an evolutionary advantage.
%
However, information processing in the DMN has also repeatedly
been shown to directly impact human behavior. Goal-directed task performance
improved with decreased activity in default mode areas \citep{weiss2006}
and increased DMN activity was linked to more task-independent,
yet sometimes useful thoughts
\citep{mason2007, seli2016mind}.
%
Gaining insight into the overarching DMN function is challenging because
this network may
simultaneously control perception-action cycles and
maintain baseline contemplations
across time, space, and content domains.

Consider an agent faced with the choice of the next action
and being guided by rich
reenactment of really-happened, hypothetically conceived, and
predicted future events to perfect ongoing behavioral performance.
A particularly attractive computational framework
to describe, quantify, and predict autonomously acting systems like the brain
can be the combination of control theory and reinforcement learning.
It is known that, the more the external world is predictable,
the more mental activity becomes detached from the actual sensory environment
\citep{antrobus1966studies, pope1978regulation}.
Conversely,
the more the ongoing executed task is unknown and unpracticed,
the less stimulus-independent thoughts occur
\citep{filler1973daydreaming, teasdale1995stimulus}.
These ``offline'' processes may however contribute to optimized control of the organism. \textbf{(XXX Guillaume: What about the idea of model free control being also remote of environment? - danilo: hmm, let's keeping thinking about this)}
% We formalize
% a policy matrix to capture the space of possible actions that can be performed
% on the environment given the current state. A value matrix
% maps expected rewards to environmental objects and events.
% Switching between states reduces to a sequential processing model.
% In a model-based reinforcement learning (RL) approach,
% we try to understand the environment always better
% starting, at birth, with an environment that is completely unknown.
Informed by outcomes of performed action,
the DMN dynamics are constantly adapted in feedback loops
shaped by prediction error.
A DMN framework incorporating RL can naturally embed human behavior
into the tension between exploitative action with immediate gains and
exploratory action with longer-term reward schedules.

@@

\section{Known neurobiological properties of the default mode network}
We begin by a deconstruction of the DMN based on experimental findings in neuroscience. This walkthrough across each individual network node will describe their mechanistic relevance, paving the way for their functional interpretation within our computational account in the next section.

@@

The PMC myelinates relatively late during postnatal development in monkeys
\citep{goldman1987development}, which is generally considered to
be a sign of phylogenetic sophistication \citep{flechsig1920}.
Physiological and disturbed metabolic fluctuations in the
human PMC have been repeatedly related to
phenomena of changed conscious awareness \citep{cavanna2006precuneus},
including anesthesia, sleep, and coma.
%
Indeed,
the PMC has long been speculated to reflect constant computation of
the environmental statistics and its internal representation
as an "inner minds eye" \citep{cavanna2006precuneus, leech_pcc2014}.
For instance, B\'alint's syndrome is a neurological disorder of conscious
awareness that results from damage in the bilateral parietal cortex
\citep{balint1909seelenlahmung}.
Patients are plagued by an
inability to bind various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as inability to direct action towards
currently unattended environmental objects
(i.e., optic ataxia).
This can be viewed as a high-level impairment in the gathering
of information about alternative objects (i.e., exploration) as well as
leveraging these environmental opportunities (i.e., exploitation).
Congruently,
the human PMC was coupled in two functional connectivity modalities
with the amygdala, involved in significance evaluation, and
the nucleus accumbens (NAc), involved in reward evaluation.
Specifically, among all parts of the PMC,
the human ventral posterior cingulate cortex was
most connected to the laterobasal
nuclei group of the amygdala
\citep{bzdok2015subspecialization}.
This amygdalar subregion has been proposed to
continuously scan environmental input
for biological significance assessment.

The putative role of the PMC in continuous integration of environmental relevance
and ensuing top-level control of action in the environment is supported
by a number of neuroscience experiments.
Electrophysiological recordings in animals implicated the PMC in
strategic decision making \citep{pearson2009neurons},
risk assessment \citep{mccoy2005risk},
and outcome-contingent behavioral modulation \citep{hayden2009electrophysiological},
while its retrosplenial portion was
more specifically implicated in approach-avoidance behavior
\citep{vann2009does}.
Neuron spiking activity in the PMC allowed distinguishing
whether a monkey would persue an exploratory or exploitative
behavioral strategy in a complex food foraging task \citep{pearson2009neurons}.
Further, single-cell recordings in the monkey PMC
demonstrated this brain region's sensitivity to
subjective target utility \citep{mccoy2005risk} and integration
across individual decision-making instances \citep{pearson2009neurons}.
This DMN node encoded the
preference of or aversion to options with uncertain reward outcomes
and its spiking activity was more associated with
subjectively perceived relevance of a chosen object
than by its factual value, based on an internal currency of value.
In fact, direct stimulation of PMC neurons
promoted exploratory action towards options
with unsafe reward outcomes that were previously shunned.
Graded changes in firing rates of PMC neurons
indicated changing choices in upcoming trials and their neural patterns were
distinct from spike firings that indicated choosing either option.
Similarly in humans,
the DMN has been shown to gather and integrate information
over different paragraphs in an fMRI study on auditory narratives
\citep{simony2016dynamic}.
%
The retrosplenial portion of the PMC can subserve action possibilities
and evaluation of reward contingencies by integrating these with
information from memory and different perspective frames.
Regarding memory retrieval, retrosplenial lesions have been
consistently associated with anterograde and retrograde memory impairments
of various kinds of sensory information
in rabbits and humans
\citep{vann2009does}.
Regarding perspective frames, this PMC subregion has been
proposed to mediate between the organism's egocentric
(i.e., focused on sensory input) and
allocentric (i.e., focused on world knowledge) viewpoints
in animals and humans
\citep{epstein2008parahippocampal, burgess2008spatial, valiquette2007different}.

Consequently, the PMC may contribute to overall DMN function
by monitoring the subjective outcomes
of possible decisions and integrates that information
with memory, perspective frames, and
reward schedules into higher-level strategies.
Perceived value that differs across individuals updates
statistical assessment of the environment
to predict delayed reward opportunities in the future.
In doing so, the PMC continuously adapts to changes
in both the external environment and its internal representation
that modulate strategic behavioral adjustment in volatile environments.

\subsection{The prefrontal cortex: stimulus-value association and action consideration}
Analogous to the PMC,
the dorsomedial PFC (dmPFC) is believed to subserve predominantly ambiguous processes
across time, space, and content domains in
sensory-independent top-down pathways.
In contrast to the PMC, however,
the dmPFC may be the DMN node whose function comes closest to a
``mental sketchpad'' \citep{goldman1996prefrontal}, as it
potentially subserves the de-novo generation and binding
of meaning representations instructed by stored semantics and memories.

Generally,
patients with neurological lesions in the prefrontal cortex
are known to adapt to novel situations and stimuli with difficulty
\citep{stuss1986frontal}.
The dmPFC may subserve inference, representation, and assessment
of one's own and other individuals' action considerations.
Specifically, neural activity in the human dmPFC
reflected expectations about other peoples actions and errors thereof.
dmPFC activity indeed explained the proficiency decline
of inferring other peoples' thoughts in the elderly \citep{moran2012social}.
Some dmPFC neurons in macaque monkeys exhibited a preference
for processing others', rather than own, behavior
with fine-grained adjustment of contextual circumstances \citep{yoshida2010neural}.
In fact, the topographically neighboring dorsal anterior cingulate cortex
has also been linked to computing values and efforts of
persisting a behavioral plan versus switching the
environmental context in several lesion studies \citep{kolling2016value}.
%
Such highly abstract neural computations necessarily rely on the
generation of probabilistic internal information drawing from
episodic memory retrieval, generative construction processes,
and explicit knowledge of the external world.
%
According to computational neuroimaging experiments,
the dmPFC activity preferentially models stimulus-value associations of
possible actions that are not actually executed,
whereas the vmPFC activity models value in the environment that leads performed actions
\citep{nicolle2012agent}.

In contrast to the dorsomedial prefrontal DMN,
the ventromedial portion of the prefrontal cortex (vmPFC) probably subserves
less ambiguous subjective-value-related evaluative processes and
reward-informed risk estimation of self-relevant environmental stimuli.
This DMN node is more closely associated with
orchestrating adaptive behavior by bottom-up-driven
processing of “what matters now”,
probably drawing on model-based value representations
\citep{doherty2015structure}.
Quantative lesion findings across 344 human individuals confirmed
a gross impairment in value-based decision making
\citep{glascher2012lesion}.
Indeed,
the vmPFC is preferentially connected with reward-related and limbic areas.
The vmPFC has been observed to have monosynaptical connections
with the NAc
in axonal tracing studies in monkeys \citep{haber1995orbital}.
Additionally, the gray-matter volume of the vmPFC and NAc
correlated with indices of value-guided behavior and reward attitudes
\citep{lebreton2009automatic}.
NAc activity is thought to reflect reward prediction signals from dopaminergic pathways
\citep{schultz1998predictive}
that not only modulate motivated behavior towards basic survival needs but also
subserve RL in humans more broadly \citep{doherty2015structure}.
This is consistent with diffusion MRI tractography in humans and monkeys
\citep{croxson2005quantitative} that
quantified this brain regions to
be substantially more likely connected to the vmPFC than dmPFC in both species.
Two functional connectivity modalities in humans strongly connected
the vmPFC with the NAc, hippocampus (HC),
and posteromedial cortex.
%
The vmPFC is often proposed to be involved in (external) emotional
reactions and own (visceral) arousal.
Real or imaginged bodily states could be mapped in the vmPFC
as a bioregulatory disposition governing cognition
and decision making \citep{damasio1996somatic}
In neuroeconomic studies of human decision making,
the vmPFC consistently reflects an individual’s subjective
valuation
\citep{behrens2008associative}.
This may be why performance within and across participants
was related to state encoding in vmPFC.
Such a ``cognitive map'' of action space was argued to encode
the current task state even when states are unobservable from sensory input,
which was shown to be critical for behavior \citep{Schuck2016}.
Additionally,
independent whole-brain analyses from structural
neuroimaging studies related the gray-matter volume of the vmPFC,
more consistently than any other
brain region, to indices of
social competence and social network complexity, which are
among the most complicated decision that humans and monkeys take
\citep{behrens2009computation}.
%

\subsection{The hippocampus: memory, space, and experience replay}
The midline core of the DMN is probably in constant communication
with the HC in the medial temporal lobe \citep{vincet2006}.
The HC is long known to be involved in memory and spatial navigation in animals and humans.

Its highly recursive anatomical architecture
may be specifically designed to allow reconstructing
entire episodes of experience from memory fragments.
%
While the HC
is traditionally believed to allow remembering the past,
there is now increasing evidence for a role
in constructing mental models in general
\citep{maguire2016, schacter2007remembering, gelbard2008internally, Javadi2017}.
Indeed,
hippocampal damage is
not only associated with an impairment in reexperiencing the past (i.e., amnesia),
but also thinking about one's own future and
imagining new fictitious experiences more broadly \citep{hassabis2007patients}.
Mental scenes created by hippocampus-lesioned patients exposed a lack of
spatial integrity, richness in detail, and overall coherence.
%
Single-cell recordings in the animal hippocampus revealed
some constantly active neuronal ensembles whose firing coincided with
distinct locations in space while the animal navigated through its surroundings.
London taxi drivers, individuals with high performance in spatial navigation,
were shown to exhibit increased grey matter volume in the
hippocampus \citep{maguire2000navigation}.
Indeed, place-specific neuronal populations in the hippocampus
spike one after another when an animal is choosing between alternative
paths \citep{johnson2007neural}; and
these neuronal patterns appear to also indicate upcoming behavior
\citep{pfeiffer2013hippocampal}.

But encoding and generative reconstruction in the hippocampus extends
beyond mere spatial knowledge of the environment.
Based on large-scale recordings of hippocampal neuronal populations,
complex spiking patterns can be followed across extended periods including
their modification of input-free self-generated patterns
after environmental events \citep{buzsaki2004large}.
Specific spiking sequences, that were elicited in experimental task conditions,
have been shown to be reenacted spontaneously during
quiet wakefulness and sleep \citep{hartley2014space, o2010play}.
Moreover, spike sequences measured in hippocampal place cells of rats
featured reoccurrence directly after experimental trials
as well as directly before upcoming experimental trials \citep{diba2007forward}.
Such hippocampal ensemble burts during rest and sleep
have been proposed to be critical in communicating local information
to the neocortex for long-term storage, potentially also in the nodes of the DMN.
These hippocampus-subserved mechanisms
probably make important contributions to the
recollection of autobiographical memory episodes and other
reexperienced or newly generated mental scenarios
\citep{hassabis2007patients}.

The HC thus orchestrates elements of experienced environmental aspects for
consolidations based on reenactment and for integration into
rich mental scene construction \citep{deuker2016event, bird2010establishing}.
In this way, the HC may even influence
ongoing perception in the current environment
\citep{maguire2016}.
% (TODO: relate to self-play in deepmind Go).
% (TODO: Kai had an interesting idea: the deja-vu effect
% could be when you already accidentally perturbed past
% such that it really occures in reality later.)

\subsection{The right and left TPJ: prediction error signaling and world semantics}
While the DMN emerges with its midline structures early in human development,
the right and left TPJ may become fully integrated into this canonical
network only after birth \citep{doria2010}.
The TPJs are known to have hemispheric differences
according to their cytoarchitectonic borders and gyrification pattern
\citep{seghier2013angular}.
Neuroscientific investigations on hemispheric functional specialization
have highlighted the right versus left cerebral hemisphere as dominant for
attention versus language functions.

>>>>

\begin{abstract}
The default mode network (DMN) is believed to subserve the human
baseline mental activity.
%
Many research streams converge to its likely role in evolutionarily adaptive
envisioning of mental scenarios to simulate the future.
Such functional roles become more likely due to
this brain network's highest energy consumption in the brain and
its intimate coupling with conscious awareness
\textbf{(XXX Guillaume: It escaladed quickly.)}.
%
The present concept paper proposes a \textit{process model}
in trying to explain \textit{how} the DMN may actually implement continuous
assessment and prediction of the environment to guide action choices.
DMN function is recast in the mathematical terms of
control theory and reinforcement learning based on Markov decision processes.
We argue that our formal account of DMN function
naturally accommodates
previously proposed cognitive accounts as special cases and
parsimoneously explains
experimental findings in animals and humans.
%
A process models for the neural implemention
of continuous future predictions probably realized in the DMN
could offer insight into the statistical mechanisms underlying human behavior.
\end{abstract}

@@

\section{Introduction}
%
When left unperturbed, the human brain is not at rest.
Rather, the brain continues to metabolize large quantities of
oxygen and glucose energy to maintain
potentially purposeful neuronal computation
in the absence of behavioral focus
\citep{kenet2003spontaneously, fiser2004small}.
The baseline energy demand is subject to surprisingly little modulations
when mentally processing environmental challenges
\citep{raichle2001pnas}. \textbf{(XXX Guillaume: What about the energy consumption compare to sleep? Danilo: I'd love to cite that
but I am not aware of such studies)}
This dark matter of brain physiology begs the question
of underlying biological purpose
\citep{raichle2006brain}.
What has early been described as the "stream of consciousness"
in psychology \citep{james1890principles}
found a potential neurobiological manifestation
in the so-called "default mode network" (DMN)
\citep{shul1997, raichle2001pnas}.
This set of some of the highest regions in
the human association cortex \citep{mesulam1998sensation, margulies2016situating}
consistently
increased in neural activity during
unfocused everyday mind wandering \citep{raichle2001pnas}.
The DMN may exert higher-order control on human behavior
following the notion of "a hierarchy of brain systems with
the DMN at the top and the salience and dorsal attention systems
at intermediate levels, above thalamic and unimodal sensory
cortex" \citep{carhart2010default}.
In the beginning of the 21st century,
brain imaging may have been first technique
to allow for the discovery of a unique brain network that subserves
baseline mental activities
\citep{raichle2001pnas, bzdok2015resting}.

\subsection{Default mode network: higher-order control of the organism}
The human DMN accounts for a large part of the brain's baseline activity
and typically decreases in activity during psychological experiments.
The standard mode of
neural information maintenance and manipulation has
been argued to mediate evolutionarily conserved functions
\citep{brown1914nature, binder1999conceptual, buzsaki2006rhythms}.
This is all the more likely because the DMN contains hotspots
of highest energy consumption in the entire central nervous system
\citep{raichle2001pnas}.
DMN activity also persists to a substantial degree during
sleep and under anesthesia \citep{randy2008}.
Today, many authors believe that the DMN implements
probabilistic estimation of past, hypothetical, and
future events
\citep{fox2005, hassabis2007patients, schacter2007remembering, binder2009,
randy2008}.
This brain network
might have emerged to continuously predict environmental events using
mental imagery as an evolutionary advantage.
%
However, information processing in the DMN has also repeatedly
been shown to directly impact human behavior. Goal-directed task performance
improved with decreased activity in default mode areas \citep{weiss2006}
and increased DMN activity was linked to more task-independent,
yet sometimes useful thoughts
\citep{mason2007, seli2016mind}.
%
Gaining insight into the overarching DMN function is challenging because
this network appears to
simultaneously control perception-action cycles and
maintain baseline contemplations
across time, space, and content domains.

In this work,
we consider an agent faced with the choice of the next action
and being guided by rich
reenactment of really-happened, hypothetically conceived, and
predicted future events to perfect ongoing behavioral performance.
A particularly attractive computational framework
to describe, quantify, and predict autonomously acting systems like the brain
is proposed to be the combination of control theory and
reinforcement learning (RL).
It is known that, the more the external world is predictable,
the more mental activity becomes detached from the actual sensory environment
\citep{antrobus1966studies, pope1978regulation}.
Conversely,
the more the ongoing executed task is unknown and unpracticed,
the less stimulus-independent thoughts occur
\citep{filler1973daydreaming, teasdale1995stimulus, christoff2016mind}.
These ``offline'' processes may however contribute to optimized control of the organism. \textbf{(XXX Guillaume: What about the idea of model free control being also remote of environment? - danilo: hmm, let's keeping thinking about this)}
% We formalize
% a policy matrix to capture the space of possible actions that can be performed
% on the environment given the current state. A value matrix
% maps expected rewards to environmental objects and events.
% Switching between states reduces to a sequential processing model.
% In a model-based reinforcement learning (RL) approach,
% we try to understand the environment always better
% starting, at birth, with an environment that is completely unknown.
Informed by outcomes of performed action,
the DMN dynamics are constantly adapted in feedback loops
shaped by prediction error.
A DMN framework incorporating RL can naturally embed human behavior
into the tension between exploitative action with immediate gains and
exploratory action with longer-term reward schedules.

@@

\section{Known neurobiological properties of the default mode network}
We begin by a neurobiological deconstruction of the DMN
based on experimental findings in neuroscience. This walkthrough across individual network nodes will describe their mechanistic relevance, paving the way for their functional interpretation within our computational account in the next section.

@@

The PMC myelinates relatively late during postnatal development in monkeys
\citep{goldman1987development}, which is generally considered to
be a sign of evolutionary sophistication \citep{flechsig1920}.
Physiological and disturbed metabolic fluctuations in the
human PMC have been repeatedly related to
phenomena of changed conscious awareness \citep{cavanna2006precuneus},
including anesthesia, sleep, and coma.
%
Indeed,
the PMC has long been speculated to reflect constant computation of
environmental statistics and its internal representation
as an "inner minds eye" \citep{cavanna2006precuneus, leech_pcc2014}.
For instance, B\'alint's syndrome is a neurological disorder of conscious
awareness that results from damage in the parietal cortex
\citep{balint1909seelenlahmung}.
Patients are plagued by an
inability to bind various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as inability to direct action towards
currently unattended environmental objects
(i.e., optic ataxia).
This can be viewed as a high-level impairment in the gathering
of information about alternative objects (i.e., exploration) as well as
leveraging these environmental opportunities (i.e., exploitation).
Congruently,
the human PMC was coupled in two functional connectivity modalities
with the amygdala, involved in significance evaluation, and
the nucleus accumbens (NAc), involved in reward evaluation.
Specifically, among all parts of the PMC,
the ventral posterior cingulate cortex was
most connected to the laterobasal
nuclei group of the amygdala
\citep{bzdok2015subspecialization}.
This amygdalar subregion has been proposed to
continuously scan environmental input
for biological significance assessment.

The putative role of the PMC in continuous integration of environmental relevance
and ensuing top-level control of action on the environment is supported
by a number of neuroscience experiments.
Electrophysiological recordings in animals implicated the PMC in
strategic decision making \citep{pearson2009neurons},
risk assessment \citep{mccoy2005risk},
and outcome-contingent behavioral modulation \citep{hayden2009electrophysiological},
while its retrosplenial portion was
more specifically implicated in approach-avoidance behavior
\citep{vann2009does}.
Neuron spiking activity in the PMC allowed distinguishing
whether a monkey would persue an exploratory or exploitative
behavioral strategy in a complex food foraging task \citep{pearson2009neurons}.
Further, single-cell recordings in the monkey PMC
demonstrated this brain region's sensitivity to
subjective target utility \citep{mccoy2005risk} and integration
across individual decision-making instances \citep{pearson2009neurons}.
This DMN node encoded the
preference of or aversion to options with uncertain reward outcomes
and its spiking activity was more associated with
subjectively perceived relevance of a chosen object
than by its factual value, based on an internal currency of value.
In fact, direct stimulation of PMC neurons
promoted exploratory action towards options
with unsafe reward outcomes that were previously shunned.
Graded changes in firing rates of PMC neurons
indicated changing choices in upcoming trials and their neural patterns were
distinct from spike firings that indicated choosing either option.
Similarly in humans,
the DMN has been shown to gather and integrate information
over different paragraphs in an fMRI study on auditory narratives
\citep{simony2016dynamic}.

The retrosplenial portion of the PMC can subserve action possibilities
and evaluation of reward contingencies by integrating these with
information from memory and different perspective frames.
Regarding memory retrieval, retrosplenial lesions have been
consistently associated with anterograde and retrograde memory impairments
of various kinds of sensory information
in rabbits and humans
\citep{vann2009does}.
Regarding perspective frames, this PMC subregion has been
proposed to mediate between the organism's egocentric
(i.e., focused on sensory input) and
allocentric (i.e., focused on world knowledge) viewpoints
in animals and humans
\citep{epstein2008parahippocampal, burgess2008spatial, valiquette2007different}.

Consequently, the PMC may contribute to overall DMN function
by monitoring the subjective outcomes
of possible decisions and integrating that information
with memory, perspective frames, and
reward schedules into higher-level strategies.
Perceived value, that differs across individuals, may update
statistical assessment of the environment
to predict delayed reward opportunities in the future.
In doing so, the PMC continuously adapts to changes
in both the external environment and its internal representation
to enable strategic behavior in volatile environments.

\subsection{The prefrontal cortex: stimulus-value association and action consideration}
Analogous to the PMC,
the dorsomedial PFC (dmPFC) of the DMN is believed to subserve
supra-sensory processes
across time, space, and content domains to
exert top-down control.
In comparison to the PMC, however,
the dmPFC may be closer to a
``mental sketchpad'' \citep{goldman1996prefrontal}, as it
potentially subserves the de-novo generation and binding
of meaning representations instructed by stored semantics and memories.

Generally,
patients with neurological lesions in the prefrontal cortex
are known to struggle with
adaption to novel situations and stimuli
\citep{stuss1986frontal}.
The dmPFC may subserve inference, representation, and assessment
of one's own and other individuals' action considerations.
Specifically, neural activity in the human dmPFC
reflected expectations about other peoples' actions and
errors in these predictions.
Neural activity in the dmPFC indeed explained the proficiency decline
of inferring other peoples' thoughts in the elderly \citep{moran2012social}.
Some dmPFC neurons in macaque monkeys exhibited a preference
for processing others', rather than own, behavior
with fine-grained adjustment of contextual circumstances \citep{yoshida2010neural}.
In fact, the topographically neighboring dorsal anterior cingulate cortex
has also been linked to computing values and efforts of
persisting a behavioral plan versus switching the
environmental context in several lesion studies \citep{kolling2016value}.
%
Such highly abstract neural computations necessarily rely on the
generation of probabilistic internal information drawing from
episodic memory retrieval, generative construction processes,
and explicit knowledge of the external world.
%
According to computational neuroimaging experiments,
dmPFC activity preferentially models stimulus-value associations of
possible actions that are not actually executed,
whereas the vmPFC activity models value in the environment that leads performed actions
\citep{nicolle2012agent}.

Comparing to the dorsomedial prefrontal DMN,
the ventromedial portion of the prefrontal cortex (vmPFC) probably subserves
less ambiguous subjective-value-related evaluative processes and
reward-informed risk estimation of self-relevant environmental stimuli.
This DMN node is more closely associated with
orchestrating adaptive behavior by bottom-up-driven
processing of “what matters now”,
probably drawing on sophisticated value representations
\citep{doherty2015structure}.
Quantitative lesion findings across 344 human individuals confirmed
a substantial impairment in value-based decision making
\citep{glascher2012lesion}.
Indeed,
the vmPFC is preferentially connected with reward-related and limbic areas.
The vmPFC is well known to have monosynaptical connections
with the NAc
in axonal tracing studies in monkeys \citep{haber1995orbital}.
Congruently, the gray-matter volume of the vmPFC and NAc
correlated with indices of value-guided behavior and reward attitudes
in humans
\citep{lebreton2009automatic}.
NAc activity is thought to reflect reward prediction signals from dopaminergic pathways
\citep{schultz1998predictive}
that not only modulate motivated behavior towards basic survival needs but also
subserve RL in humans more broadly \citep{doherty2015structure}.
This is consistent with diffusion MRI tractography in humans and monkeys
\citep{croxson2005quantitative} that
quantified the NAc to
be substantially more likely connected to the vmPFC than dmPFC in both species.
Two functional connectivity modalities in humans strongly connected
the vmPFC with the NAc, hippocampus (HC),
and PMC.
%
The vmPFC is often proposed to be involved in (external) emotional
reactions and own (visceral) arousal.
Real or imaginged bodily states could be mapped in the vmPFC
as a bioregulatory disposition governing cognition
and decision making \citep{damasio1996somatic}
In neuroeconomic studies of human decision making,
the vmPFC consistently reflects an individual’s subjective
valuation
\citep{behrens2008associative}.
This may be why performance within and across participants
was related to state encoding in vmPFC.
Such a ``cognitive map'' of the action space was argued to encode
the current task state even when states are unobservable from sensory input,
which was shown to be critical for behavior \citep{Schuck2016}.
Additionally,
independent whole-brain analyses from structural
neuroimaging studies related the gray-matter volume of the vmPFC,
more consistently than any other
brain region, to indices of
social competence and social network complexity, which are probably
among the most complicated decision that humans and monkeys take
\citep{behrens2009computation}.

\subsection{The hippocampus: memory, space, and experience replay}
The DMN midline is probably in constant communication
with the HC in the medial temporal lobe \citep{vincet2006}.
The HC is long known to be involved in
memory operations and spatial navigation in animals and humans.

While the HC
is traditionally believed to allow remembering the past,
there is now increasing evidence for a role
in constructing mental models in general
\citep{maguire2016, schacter2007remembering, gelbard2008internally, Javadi2017}.
Its highly recursive anatomical architecture
may be specifically designed to allow reconstructing
entire episodes of experience from memory fragments.
Indeed,
hippocampal damage is
not only associated with an impairment in reexperiencing the past (i.e., amnesia),
but also thinking about one's own future and
imagining new fictitious experiences more broadly \citep{hassabis2007patients}.
Mental scenes created by hippocampus-lesioned patients exposed a lack of
spatial integrity, richness in detail, and overall coherence.
%
Single-cell recordings in the animal hippocampus revealed
some constantly active neuronal ensembles whose firing coincided with
distinct locations in space during environmental navigation.
London taxi drivers, individuals with high performance in spatial navigation,
were shown to exhibit increased gray-matter volume in the
hippocampus \citep{maguire2000navigation}.
Indeed, place-specific neuronal populations in the hippocampus
spike one after another when an animal is choosing between alternative
paths \citep{johnson2007neural}.
These neuronal patterns appear to also indicate upcoming behavior
\citep{pfeiffer2013hippocampal}.

But encoding and generative reconstruction in the hippocampus extends
beyond mere spatial knowledge of the environment.
Based on large-scale recordings of hippocampal neuronal populations,
complex spiking patterns can be followed across extended periods including
their modification of input-free self-generated patterns
after environmental events \citep{buzsaki2004large}.
Specific spiking sequences, that were elicited in experimental task conditions,
have been shown to be reenacted spontaneously during
quiet wakefulness and sleep \citep{hartley2014space, o2010play}.
Moreover, spike sequences measured in hippocampal place cells of rats
featured re-occurrence directly after experimental trials
as well as directly before upcoming experimental trials \citep{diba2007forward}.
Such hippocampal ensembles burt during rest and sleep
have been proposed to be critical in communicating local information
to the neocortex for long-term storage, potentially also in the nodes of the DMN.
These hippocampus-subserved mechanisms
probably contribute to
recollecting autobiographical memory episodes, as well as
reexperiencing or newly generating mental scenarios
\citep{hassabis2007patients}.

The HC thus orchestrates experience of environmental aspects for
consolidations based on reenactment and for integration into
rich mental scene construction \citep{deuker2016event, bird2010establishing}.
In this way, the HC may even influence
ongoing perception in the current environment
\citep{maguire2016}.
% (TODO: relate to self-play in deepmind Go).
% (TODO: Kai had an interesting idea: the deja-vu effect
% could be when you already accidentally perturbed past
% such that it really occures in reality later.)

\subsection{The right and left TPJ: prediction error signaling and world semantics}
While the DMN emerges with its midline structures early in human development,
the right and left TPJ may become fully integrated into this canonical
network only after birth \citep{doria2010}.
The TPJs are known to exhibit hemispheric differences
according to their cytoarchitectonic borders and gyrification pattern
\citep{seghier2013angular}.
Neuroscientific investigations on hemispheric functional specialization
have highlighted the right versus left cerebral hemisphere as dominant for
attention versus language functions.

The TPJ in the right hemipshere (RTPJ) is central for
action initiation during externally structured tasks and
sensorimotor control by integrating supra-modal stimulus-guided attention
\citep{corbetta2002control}.
Involvement of this DMN node was repeatedly reported in
multi-step action execution \citep{hartmann2005takes},
visuo-proprioceptive conflict \citep{Balslev2005} and
multi-modal detection of sensory changes across
visual, auditory, or tactile stimulation in a multi-modal fMRI study
\citep{downar2000multimodal}.
In humans, direct electrical stimulation of the
RTPJ during neurosurgery was associated with altered perception
and stimulus awareness \citep{blanke2002neuropsychology}.
%
Importantly, the RTPJ has been shown to be intimately related to
supra-modal prediction and error signaling.
It was argued that the RTPJ encodes actions and ensuing outcomes
without necessarily relating those to outcome value
\citep{liljeholm2013neural, hamilton2008action,
jakobs2009effects}.
Neural activity in the RTPJ has been argued to be responsible
for stimulus-driven attentional reallocation to
salient and surprising sources of information
as a ‘‘circuit breaker’’ that recalibrates control and maintenance systems
\citep{bzdok2013tpj, corbettashul2008}.
% (Bzdok et al., 2013; Corbetta et al., 2008).
Indeed, patients with right TPJ damage have particular difficulties
with multistep actions \citep{hartmann2005takes}.
In the face of large discrepancies between actual and previously predicted
environmental events the RTPJ acts a potential switch between
externally-oriented mind sets focussed on the
sensory world and internally-oriented mind sets focussed
on self-relevant mental scene construction.
Transient RTPJ in humans for instance disrupted the
impact of predicted intentions of other individuals
\citep{young2010disruption},
a capacity believed to be subserved by the DMN.
The RTPJ might hence be an important relay that shifts away
from the ‘‘internally directed’’ baseline processes
to, instead, deal with environmental objects and events.
% (TODO: slow endogeneous changes in state transitions by the policy and value matrices)
